Submission and Formatting Instructions for the Twenty-Eighth \\ International Conference on Machine Learning (ICML 2012)

### 2012-02-22 klein2012/main.tex
voila

The reward function $R$ is the unknown of the problem. We can suppose that it is possible to dispose of $p$ features $\psi_{1\leq i \leq p}$, which are functions from $S$ to $\mathbb{R}$, such  that:

@@

a^*_i &=& \arg\max_{a\in A}(\theta ^T \mu^E(s_i,a) + l(s_i,a)),\\

@@

\theta_{t+1} &\leftarrow& \theta_t -\alpha_t\frac{g_{\theta}}{\|g_{\theta}\|_2}.

@@

%Dans le cas idéal où le modèle est connu, le calcul exact de $\mu^E$ peut être effectué grâce aux algorithmes de programmation dynamique. Dans le cas plus réaliste où les probabilités de transitions sont inconnues, on peut, si l'on dispose d'un simulateur et qu'il est possible d'interroger l'expert en un grand nombre d'états arbitraires, utiliser une simple approximation de Monte-Carlo où pour approximer $\mu^E(s,a)$ (définition Eq. \eqref{Qmu.eqn}) l'on utilise
In the ideal case where a model of the environnement is known (\textit{id est} the transition probabilities are given), the exact calculus of $\mu^E$ can be executed thanks to dynamic programmation algorithms. In a more realistic case, we can use a simple Monte-Carlo estimation if we dispose of a simulator and if it is possible to ask the expert what action has to be chosen in an important number of arbitrary states.The estimation of $\mu^E(s,a)$ (definition Eq. \eqref{Qmu.eqn}) is done over $N$ trajectories of length $\{2k_i+1\}_{1\leq i \leq N}$, $\{s^i_0,a^i_0,\dots,s^i_{k_i-1},a^i_{k_i-1},s^i_{k_i}\}_{1\leq i \leq N}$, by the following formula: \begin{equation}
\hat\mu^E(s,a) = {1\over N}\sum_{i=1}^{N}\sum^{k_i}_{t=0}\gamma^t\psi(s^i_t),
\end{equation}
where $s^i_0=s, a^i_0=a$ and $s^i_{t+1}$ is distributed according to $P_{s^i_t\pi_E(s^i_t)}$.

@@

In this section we are going to prove that our algorithm (see \eqref{algorithm}) is convergent which means that $\tilde{\theta}=\lim_{t\rightarrow+\infty}\theta_t$ is well defined and that $\inf_{\theta\in\mathbb{R}^p}J_N(\theta)=J_N(\tilde{\theta})$ which means that the algorithm reaches the goal of minimizing $J_N$. Moreover we will determine the quality of our algorithm by giving an upper bound to the term $E_{s\sim\nu}[V^{\pi_E}(s)-Q^{\pi_E}(s,\pi_A(s))]$ where $E_{s\sim\nu}$ is the expectation over the stationary probability distribution of states $\nu$ and $\pi_A$ is the policy of the apprentice.

@@

Let $N\in\mathbb{N}^*$ and $A$ a finite set of cardinal $\card{A}$.
Let $\{f_{i,a}\}_{1\leq i \leq N,a\in A}$ and $\{l_{i,a}\}_{1\leq i \leq N,a\in A}$ be finite sets of cardinal $N\card{A}$ which respectively each element $f_{i,a}\in\mathbb{R}^p$ and $l_{i,a}\in\mathbb{R_+}$  where $p\in\mathbb{N}^*$. $\mathbb{R}^p$ is an euclidian space with the scalar product $<.,.>$ and the associate euclidian norm $\|\|_2$. 
We define then the function $f_N$ from $\mathbb{R}^p$ to $\mathbb{R}$ such that $\forall \theta\in\mathbb{R}^p, f_N(\theta)=\frac{1}{N}\sum_{i=1}^{i=N}\max_{a\in A}(\theta^Tf_{i,a}+l_{i,a})$. It is clear, as $A$ is finite, that $f_N$ is convex, proper, closed, continuous and with finite values.
Moreover one of its sub-gradient in $\theta\in\mathbb{R}^p$ is $g_{\theta}=\frac{1}{N}\sum_{i=1}^{i=N}f_{i,a_i^*}$ where $a_i^*\in\argmax_{a \in A}(\theta^Tf_{i,a}+l_{i,a})$. We also have $g_{\theta}\in\{\frac{1}{N}\sum_{i=1}^{i=N}f_{i,a}\}_{1\leq i \leq N,a\in A}=G$, $G$ is a finite set . Let note $g_{min}$ one element of $\argmin_{G\backslash{0}}(\frac{1}{N}\|\sum_{i=1}^{i=N}f_{i,a}\|_2)$, and
$\overline{f}_N=\inf_{\theta\in\mathbb{R}^p}f_N(\theta)$. Finally, we suppose $f_N(0)\geq \overline{f}_N\geq0$. 

@@

>>>>

The reward function $R$ is the unknown of the problem. We can suppose that it is possible to dispose of $p$ features via $\psi: S\rightarrow \mathbb{R}^p$, which is a functions from $S$ to $\mathbb{R}^p$ and each of the components of $\psi$: $\{\psi_i\}_{1\leq i \leq p}$ is one feature, such that:

@@

a^*_i &\in& \arg\max_{a\in A}(\theta ^T \mu^E(s_i,a) + l(s_i,a)),\\

@@

\theta_{0}&=&0,\theta_{t+1}=\theta_t -\alpha_t\frac{g_{\theta}}{\|g_{\theta}\|_2},\forall t\in\mathbb{N}^*,\\
\sum_{t\geq0}\alpha_t&>&+\infty,\text{ }\sum_{t\geq0}\alpha_t^2<+\infty.

@@

%Dans le cas idéal où le modèle est connu, le calcul exact de $\mu^E$ peut être effectué grâce aux algorithmes de programmation dynamique. Dans le cas plus réaliste où les probabilités de transitions sont inconnues, on peut, si l'on dispose d'un simulateur et qu'il est possible d'interroger l'expert en un grand nombre d'états arbitraires, utiliser une simple approximation de Monte-Carlo où pour approximer $\mu^E(s,a)$ (définition Eq. \eqref{Qmu.eqn}) l'on utilise
In the ideal case where a model of the environnement is known (\textit{id est} the transition probabilities are given), the exact calculus of $\mu^E$ can be executed thanks to dynamic programmation algorithms. In a more realistic case, we can use a simple Monte-Carlo estimation if we dispose of a simulator and if it is possible to ask the expert what action has to be chosen in an important number of arbitrary states.The estimation of $\mu^E(s,a)$ (definition Eq. \eqref{Qmu.eqn}) is done over $N$ trajectories of length $\{2k_i+1\}_{1\leq i \leq N}$, $\{s^i_0,a^i_0,\dots,s^i_{k_i-1},a^i_{k_i-1},s^i_{k_i}\}_{1\leq i \leq N}$, by the following formula:
\hat\mu^E(s,a) = {1\over N}\sum_{i=1}^{N}\sum^{k_i}_{t=0}\gamma^t\psi(s^i_t),
where $s^i_0=s, a^i_0=a$ and $s^i_{t+1}$ is distributed according to $P_{s^i_t\pi_E(s^i_t)}$.

@@

In this section we are going to prove that our algorithm (see \eqref{algorithm}) is convergent, when we suppose the existence of a minimum of $J_N$,  which means that $\tilde{\theta}=\lim_{t\rightarrow+\infty}\theta_t$
is well defined and that $\inf_{\theta\in\mathbb{R}^p}J_N(\theta)=J_N(\tilde{\theta})$ which means that the algorithm reaches the goal of minimizing $J_N$.
Moreover we will determine the quality of our algorithm by giving an upper bound to the term $E_{s\sim\nu}[V^{\pi_E}(s)-Q^{\pi_E}(s,\pi_A(s))]$ where $E_{s\sim\nu}$ is the expectation
over the stationary probability distribution of states $\nu$ and $\pi_A$ is the policy of the apprentice when a specific condition is fulfilled.

@@

$\mathbb{R}^p$ is an euclidian space with the canonical scalar product $<.,.>$ and the euclidian norm $\|\|_2$.
Let $N\in\mathbb{N}^*$ and $A$ the finite set of actions of cardinal $\card{A}$. We dispose of the estimations  $\{\hat{\mu}^E_{ia}\}_{1\leq i \leq N,a\in A}$ where $\forall 1\leq i \leq N, \forall a\in A,\hat{\mu}^E_{ia}\in\mathbb{R}^p$.
We define the loss function such that: $\forall 1\leq i \leq N,\forall a\in A\backslash{a_i}, l(s_i,a)>l(s_i,a_i)\geq0$. We have already defined the empirical risk function which is a function from $\mathbb{R}^p$ to $\mathbb{R}$:
$\forall \theta\in\mathbb{R}^p, J_N(\theta)= {1\over N} \sum_{i=1}^N\left(\max_{a\in A}(\theta ^T \hat{\mu}^E_{s_ia} + l(s_i,a)) - \theta ^T \hat{\mu}^E_{s_ia_i} \right)$. It is clear, as $A$ is finite, that it is a convex, proper, closed and continuous function
with finite values. Moreover one sub-gradient of $J_N$ in $\theta\in\mathbb{R}^p$ is $g_{\theta}={1\over N}\sum_{i=1}^N\left(\hat{\mu}^E_{s_ia^*_i} - \hat{\mu}^E_{s_ia_i}\right)$, where $a^*_i\in\arg\max_{a\in A}(\theta ^T \hat{\mu}^E_{s_ia} + l(s_i,a))$.
We can then use the result from the article \cite{correa1993convergence} section 5 which allows to say, with our notations, that the sequence $\theta_0=0,\forall t\in \mathbb{N}^*,\theta_{t+1}=\theta_t-\alpha_t\frac{g_{\theta_t}}{\|g_{\theta_t}\|_2}$ is convergent
if there exists a minimum to $J_N$, and that $\liminf_{t\rightarrow +\infty}J_N(\theta_t)=\inf_{\theta\in\mathbb{R}^p}J_N(\theta)$ under the conditions: $\sum_{t\geq0}\alpha_t>+\infty,\text{ }\sum_{t\geq0}\alpha_t^2<+\infty$. For a discussion on the existence
of the minimum, one can report to the article \cite{kaufmann1988existence}. Finally, one can show that if there exists $\theta\in\mathbb{R}^p$ such that $\forall 1\leq i \leq N, \theta^T\mu^E_{s_ia_i}>\max_{a\in A \backslash a_i}\theta^T\mu^E_{s_i,a}$, then there exists a minimum of $J_N$ and that all the
minimums: $\{\tilde{\theta}: \inf_{\theta\in\mathbb{R}^p}J_N(\theta)=J_N(\tilde{\theta})\}$ verify $\forall 1\leq i \leq N, \tilde{\theta}^T\mu^E_{s_ia_i}>\max_{a\in A \backslash a_i}\tilde{\theta}^T\mu^E_{s_i,a}$.

@@

Here we are going to bound the term $E_{s\sim\nu}[V^{\pi_E}(s)-Q^{\pi_E}(s,\pi_A(s))]$. First of all let us introduce the estimation errors $\{\epsilon^{\mu}_{s_ia_i}=\hat{\mu}^E_{s_ia_i}-\mu^E_{s_ia_i}\}_{1\leq i \leq N,a\in A}$ and the
approximation error $\epsilon^R$ which results from the fact that the unknown and bounded reward function $R\in\mathbb{R}^S$ may not be a linear combination of the features $\psi_{1\leq i \leq p}$. Formally, let $T$ be the tuple $\{S,A,\{P_{sa}\}_{s\in S,a\in A},\gamma\}$ which is
an MDP without reward and $T(\tilde{R})=\{S,A,\{P_{sa}\}_{s\in S,a\in A},\gamma,\tilde{R}\}$ where $\tilde{R}\in\mathbb{R}^S$. We note $C_R$ the reward-equivalence class of $R$: $C_R=\{\tilde{R}\in\mathbb{R}^S: \Pi(T(R))=\Pi(T(\tilde{R}))\}$ where $\Pi(T(\tilde{R}))$ is the set of optimal deterministic policies
of $T(\tilde{R})$. We can now give a formal definition of $\epsilon^R$: $\epsilon^R=\|R^*-\theta^{*T}\psi\|_\infty=\max_{s\in S}|R^*(s)-\theta^{*T}\psi(s)|$ where $(R^*,\theta^*)\in\argmin_{\tilde{R}\in C_R, \theta\in\mathbb{R}^p}{\|\theta^T\psi-R\|_2}$.
Now we suppose that our sub-gradient converges to $\tilde{\theta}$ and that $g_{\tilde{\theta}}={1\over N}\sum_{i=1}^N\left(\hat{\mu}^E_{s_ia^*_i} - \hat{\mu}^E_{s_ia_i}\right)=0$, where $a^*_i\in\arg\max_{a\in A}(\tilde{\theta}^T \hat{\mu}^E_{s_ia} + l(s_i,a))$.
This case happen for instance when there exists $\theta\in\mathbb{R}^p$ such that $\forall 1\leq i \leq N, \theta^T\mu^E_{s_ia_i}>\max_{a\in A \backslash a_i}\theta^T\mu^E_{s_i,a}$. The apprentice policy $\pi_A$ is an optimal deterministic policy
for the MDP $T(\tilde{\theta}\psi)$ and we can show that: $\forall 1\leq i\leq N, \pi_A(s_i)=a_i^*$. So we have: ${1\over N}\sum_{i=1}^N\left(\hat{\mu}^E_{s_i\pi_A(s_i)} - \hat{\mu}^E_{s_i\pi_E(s_i)}\right)=0$ which implies: ${1\over N}\sum_{i=1}^N(<\mu^E_{s_i\pi_A(s_i)},\theta^*>-<\mu^E_{s_i\pi_E(s_i)},\theta^*>+<\epsilon^{\mu}_{s_i\pi_A(s_i)}-\epsilon^{\mu}_{s_i\pi_E(s_i)},\theta^*>)=0$.
Here $Q^{\pi_E}$ and $V^{\pi_E}$ are defined in function of $R^*$ and not in function of $R$, we should use the notations $Q^{\pi_E}_{R^*}$ and $V^{\pi_E}_{R^*}$.
Let note the action-value errors $\{\epsilon^Q_{s_ia_i}=Q^{\pi_E}(s_i,a_i)-<\theta^*,\mu^E_{s_ia_i}>\}_{1\leq i \leq N,a\in A}$, we have $\forall 1\leq i\leq N, |\epsilon^Q_{s_ia_i}|=|E^{\pi_E}_{s_ia_i}[\sum_{t\geq0}\gamma^t(R^*(s_t)-<\theta^*,\psi(s_t)>)]|\leq\frac{\epsilon^R}{1-\gamma}$. So we have:
${1\over N}\sum_{i=1}^N(Q^{\pi_E}(s_i,\pi_E(s_i))-\Q^{\pi_E}(s_i,\pi_A(s_i))+<\epsilon^{\mu}_{s_i\pi_E(s_i)}-\epsilon^{\mu}_{s_i\pi_A(s_i)},\theta^*>-\epsilon^{Q}_{s_i\pi_E(s_i)}+\epsilon^{Q}_{s_i\pi_A(s_i)})=0$. Finally, let us define $\epsilon^V$, $V$ as Vapnick, the Vapnick-error term:
$\epsilon^V=E_{s\sim\nu}[V^{\pi_E}(s)-Q^{\pi_E}(s,\pi_A(s))]-[{1\over N}\sum_{i=1}^NQ^{\pi_E}(s_i,\pi_E(s_i))-\Q^{\pi_E}(s_i,\pi_A(s_i))]$. Here we suppose that the samples$\{s_i,a_i\}_{1\leq i \leq N,a\in A}$ are i.i.d and distributed over $\nu$. We know that the $\pi_A$ we find depends only on $\tilde{\theta}$ which depends on the samples
but as $\tilde{\theta}$ is a linear combination of $p$ parameters it implies that the space of possible $\pi_A$ has a VC-dimension $h$ such that: $h\leq p$. So, we can use a Vapnick-bound for $\epsilon^V$ with probability $1-\eta$ because $\forall s\in S, 0\leq|Q^{\pi_E}(s,\pi_E(s))-\Q^{\pi_E}(s,\pi_A(s))|\leq\frac{R^*_{max}}{1-\gamma}$, where $R^*_{max}=\max_{s\in S}R^*(s)$. 
$|\epsilon_V|\leq\frac{R^*_{max}}{1-\gamma}\epsilon(N)$ where $\epsilon(N)=\sqrt{\frac{h}{N}[\ln(\frac{2N}{h})+1]-\frac{1}{N}\ln(\frac{\eta}{4})}$. Finally by Cauchy-Schwartz, we have the final upper-bound with probability $1-\eta$:
$E_{s\sim\nu}[V^{\pi_E}(s)-Q^{\pi_E}(s,\pi_A(s))]\leq\frac{\epsilon^R}{1-\gamma}+2\|\theta^*\|_2\|\epsilon^{\mu}\|_{\infty}+\epsilon(N)$, where $\|\epsilon^{\mu}\|_{\infty}=\max_{1\leq i\leq N,a\in A}\|\epsilon^{\mu}_{s_ia}\|_2$. 
